---
title: "Hands-on ex1"
author: "Ke Ke"
format:
  html:
    code-fold: true
    code-tools: true
---

# 1. Getting Started

## 1.1 Install and launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse pakcges are installed in the computer. If yes, then they will be launched into R environment.

```{r}
pacman::p_load(tidyverse)


```

## 1.2 Checking Working Directory and Importing Data

```{r}
getwd()
```

```{r}
exam_data <- read_csv("Exam_data.csv")
exam_data
```

## 1.3 Plotting a simple bar chart

```{r}
ggplot(data = exam_data,
       aes(x=RACE)) +
  geom_bar()
```

# 2. Introducing `ggplot`

## 2.1 Comparison of Base R and ggplot plot

```{r,fig.show='hold', fig.ncol=2}
library(ggplot2)

#R graphic 
hist(exam_data$MATHS)

#ggplot
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=10, 
                 boundary = 100,
                 color="black", 
                 fill="grey") +
  ggtitle("Distribution of Maths scores")



```

::: {.callout-note collapse="true"}
## Notes:

While both approaches achieve similar results, the ggplot2 version can be easily modified and extended with additional layers, themes, and scales without fundamentally altering the underlying code structure. more suitable for complex visualization.
:::
